---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r  echo=TRUE}
activity <- read.csv("C:/Users/Abhishek/Desktop/activity.csv")

```
Exploring the dataset
```{r echo=TRUE}
head(activity)
summary(activity)

```


## What is mean total number of steps taken per day?
```{r}
agg<-aggregate(steps~date,data=activity,na.rm=TRUE,FUN=sum)
print(agg)

```
Histogram of steps
```{r}
hist(agg$steps,xlab="Steps taken in a day",main="Histogram of steps")
```
<br/>

Finding mean and median
```{r}
Mean<-mean(agg$steps)
Mean
Median<-median(agg$steps)
Median

```

## What is the average daily activity pattern?
Time series plot
```{r}
agg2<-aggregate(steps~interval,data=activity,na.rm=TRUE,FUN=mean)
plot(agg2$interval,agg2$steps,type="l",xlab="Intreval",ylab="Steps",main="Time series Plot")
```
<br/>

Finding interval corresponding to maximum steps
```{r}
Maxi<-subset(agg2,agg2$steps==(max(agg2$steps)))
Maxi$interval
```
Total number of misssing values
```{r}
a<-subset(activity,is.na(steps))
 NROW(a)
```
## Imputing missing values
 Method to fill missing values : Replace the NA values of a particular step with the average number of steps of the corresponding interval it belongs to. Replacing missing values:
```{r}
ac2<-aggregate(steps~interval,data=activity,FUN=mean)
for (i in 1:nrow(activity))
{
  if(is.na(activity$steps[i]))
  {
    activity$steps[i]<-ac2$steps[which(ac2$interval %in% activity$interval[i])]
  }
}
```
Creating new dataset : Missing values already replaced in original dataset
```{r}
Activit<-activity
```



## Are there differences in activity patterns between weekdays and weekends?
Making histogram
```{r}
agg2<-aggregate(steps~date,data=Activit,na.rm=TRUE,FUN=sum)
hist(agg2$steps,xlab="Steps taken in a day",main="Histogram of steps")
```
Calculating mean and median
```{r}
Mean<-mean(agg2$steps)
Mean
Median<-median(agg2$steps)
Median

```
 As we can see, there is a difference between the two histograms belonging to the two datasets. The frequency has increased when the missing values have been replaced. This was expected.
There is a slight change in the mean and median as well 

Making panel plot containing plots of average number of steps taken
on weekdays and weekends.
```{r}
library(ggplot2)
activity$date<-as.POSIXlt(activity$date)
activity$day <-  ifelse(weekdays(activity$date) %in% c("Saturday","Sunday"), 'weekend', 'weekday')
table(activity$day)
agg3<-aggregate(steps~interval+day,data=activity,na.rm=TRUE,FUN=mean)
ggplot(agg3,aes(x=interval,y=steps),color=day)+geom_line()+facet_wrap(~day)

```